---
output: 
  github_document
---

# Flight Analysis - R Package

[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

An R package for analyzing, forecasting, and collecting flight data and prices from Google Flights. 

**Credits:** This package is an R implementation inspired by the original Python package [google-flight-analysis](https://github.com/celebi-pkg/flight-analysis) by Kaya Celebi.

## Features

Current features include:

- Detailed scraping and querying tools for Google Flights
- Base analytical tools/methods for price forecasting/summary
- Support for multiple trip types: one-way, round-trip, chain-trip, and perfect-chain

## Installation

You can install the development version of flightanalysis from GitHub:

```{r}
# Install devtools if you haven't already
# install.packages("devtools")

# Install flightanalysis
devtools::install_github("rempsyc/flightanalysis")
```

## Usage

### Loading the Package

```{r}
library(flightanalysis)
```

### Creating Flight Queries

The main scraping function that makes up the backbone of most functionalities is `Scrape()`. It serves as a data object, preserving the flight information as well as meta-data from your query.

#### One-Way Trip

```{r}
# Create a one-way flight query
scrape <- Scrape("JFK", "IST", "2025-07-20")
scrape
```

#### Round-Trip

```{r}
# Create a round-trip flight query
scrape <- Scrape("JFK", "IST", "2025-07-20", "2025-08-20")
scrape
```

#### Chain-Trip

Chain-trips are defined as a sequence of one-way flights that have no direct relation to each other, other than being in chronological order.

```{r}
# Chain-trip format: origin, dest, date, origin, dest, date, ...
scrape <- Scrape("JFK", "IST", "2025-08-20", "RDU", "LGA", "2025-12-25", "EWR", "SFO", "2026-01-20")
scrape
```

#### Perfect-Chain

Perfect-chains are defined as a sequence of one-way flights such that the destination of the previous flight is the origin of the next, and the origin of the chain is the final destination (a cycle).

```{r}
# Perfect-chain format: origin, date, origin, date, ..., first_origin
scrape <- Scrape("JFK", "2025-09-20", "IST", "2025-09-25", "CDG", "2025-10-10", "LHR", "2025-11-01", "JFK")
scrape
```

### Scraping Data

The package now includes full web scraping functionality using **chromote**!

**Why chromote?**
- ✅ No external driver files needed (uses Chrome DevTools Protocol directly)
- ✅ More reliable - no driver version mismatches or port conflicts
- ✅ Simpler installation - just install the R package
- ✅ Works on all platforms (Windows, macOS, Linux)
- ✅ Fully headless by default

Then scrape flight data from Google Flights:

```{r}
# Create a query
scrape <- Scrape("JFK", "IST", "2025-12-20", "2026-01-05")

# Scrape the data (runs in headless mode by default)
scrape <- ScrapeObjects(scrape)

# View the scraped data
head(scrape$data)
```

The `ScrapeObjects()` function will:
1. **Run pre-flight checks** - Verify Chrome installation and internet connectivity
2. **Automatically connect to Chrome** - Using Chrome DevTools Protocol (no drivers!)
3. **Navigate to Google Flights URLs** - With proper wait times for page loading
4. **Extract flight information** - Parse prices, times, airlines, stops, and emissions
5. **Store results** - Save data in the Scrape object's `data` field
6. **Handle errors gracefully** - Provide detailed troubleshooting tips if issues occur

**Advantages over RSelenium:**

chromote eliminates common RSelenium issues:
- ❌ No more "invalid assignment for reference class field 'port'" errors
- ❌ No more driver download/installation problems
- ❌ No more version compatibility issues
- ❌ No more port conflicts
- ✅ Just install chromote and it works!

### Working with Flight Objects

```{r}
# Create Flight objects
flight1 <- Flight("2025-12-25", "JFKIST", "9:00AM", "5:00PM", 
                  "8 hr 0 min", "Nonstop", "150 kg CO2", "10% emissions", "$450")
flight2 <- Flight("2025-12-26", "ISTCDG", "10:00AM", "2:00PM", 
                  "4 hr 0 min", "Nonstop", "100 kg CO2", "5% emissions", "$300")
flight3 <- Flight("2025-12-27", "CDGJFK", "11:00AM", "1:00PM", 
                  "8 hr 0 min", "1 stop", "200 kg CO2", "15% emissions", "$500")

# Convert multiple flights to a data frame
flights <- list(flight1, flight2, flight3)
df <- flights_to_dataframe(flights)
df
```



## Trip Types

The package supports multiple trip types:

- **One-way**: Single flight from origin to destination
- **Round-trip**: Flight to destination and return
- **Chain-trip**: Sequence of unrelated one-way flights in chronological order
- **Perfect-chain**: Sequence where each destination becomes the next origin, forming a cycle

## Package Structure

```
flightanalysis/
├── R/
│   ├── flight.R             # Flight class and methods
│   ├── scrape.R             # Scrape class and methods
│   └── flightanalysis-package.R  # Package documentation
├── tests/
│   └── testthat/            # Test files
│       ├── test-flight.R
│       └── test-scrape.R
├── DESCRIPTION              # Package metadata
├── NAMESPACE               # Package exports
└── README.Rmd             # This file
```

## Differences from Python Version

This R package maintains the core functionality of the Python version but with R-specific implementations:

- Uses S3 classes instead of Python classes
- chromote instead of Python's selenium
- Base R data frames instead of pandas
- Native R date/time handling instead of Python's datetime

## Development Status

This is a complete R port of the Python package. The core data structures and API are fully implemented. Web scraping functionality uses chromote (Chrome DevTools Protocol), which provides driver-free browser automation without the configuration overhead of RSelenium.

## License

MIT License

## Original Python Package

This is a port of the [google-flight-analysis](https://github.com/celebi-pkg/flight-analysis) Python package by Kaya Celebi.

## Contributing

Contributions are welcome! Please feel free to submit a Pull Request.
