---
output: 
  github_document
---

# Flight Analysis: Find the best flights

An R package for analyzing, forecasting, and collecting flight data and prices from Google Flights. 

## Features

- Detailed scraping and querying tools for Google Flights using chromote
- Support for multiple trip types: one-way, round-trip, chain-trip, and perfect-chain
- Flexible date search across multiple airports and date ranges
- Summary tables showing prices by city and date
- Automatic identification of cheapest travel dates
- Visualization functions for price trends and best dates

## Installation

You can install the development version of flightanalysis from GitHub:

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r install, eval=FALSE}
install.packages('flightanalysis', 
  repos = c('https://rempsyc.r-universe.dev', 'https://cloud.r-project.org'))

# Or if you need the version from the last hour, install through `remotes`
# install.packages("remotes")
remotes::install_github("rempsyc/flightanalysis")
```

## Usage

### Loading the Package

```{r library, eval=TRUE}
library(flightanalysis)
```

### Creating Flight Queries and Fetching the Data

The main scraping function that makes up the backbone of most functionalities is `fa_define_query()`. It serves as a data object, preserving the flight information as well as meta-data from your query. `fa_fetch_flights()` then fetches flight information from that query. Origin and destination airports can be specified as a mix of Airport 3-letter codes, city 3-letter codes, or city names.

```{r fa_define_query, eval=TRUE}
# Round-trip
query <- fa_define_query("NYC", "London", "2025-12-20", "2026-01-05")

# Same as:
query <- fa_define_query("New York", "LON", "2025-12-20", "2026-01-05")
query

# Fetch the flight data
flights <- fa_fetch_flights(query)

# View the flight data
head(flights$data[1:11]) |>
  knitr::kable()
```

The package supports multiple trip types:

- **One-way**: `fa_define_query("JFK", "IST", "2025-07-20")`
- **Round-trip**: `fa_define_query("JFK", "IST", "2025-07-20", "2025-08-20")`
- **Chain-trip**: `fa_define_query("JFK", "IST", "2025-08-20", "RDU", "LGA", "2025-12-25")`
- **Perfect-chain**: `fa_define_query("JFK", "2025-09-20", "IST", "2025-09-25", "JFK")`

## Flexible Date Search

The package supports flexible date search across multiple airports and dates:

```{r fa_define_query_range, eval=TRUE}
# Create query objects for multiple origins and dates
queries <- fa_define_query_range(
  origin = c("BOM", "DEL"),
  dest = "JFK",
  date_min = "2025-12-18",
  date_max = "2025-12-22"
)

# Fetch all flights
flights <- fa_fetch_flights(queries)

# Create summary table (City × Date with prices)
fa_summarize_prices(flights) |>
  knitr::kable()

# Find the cheapest dates
fa_find_best_dates(
  flights, 
  n = 5,
  by = "min",
  price_max = 1400,
  max_stops = 1,
  travel_time_max = 26  # 26 hours (numeric = hours, or use "26 hr" format)
  ) |>
  knitr::kable()
```

## Visualizing Price Data

The package includes plotting functions to visualize price trends and best dates:

```{r plots, eval=TRUE}
# Plot price trends across dates
fa_plot_prices(flights, 
               title = "Flight Prices: BOM/DEL to JFK",
               size_by = "travel_time",
               annotate_col = "travel_time")

# Plot best travel dates
fa_plot_best_dates(flights)
```

## Original Python Package

**Credits:** This package is an R implementation inspired by the original Python package [google-flight-analysis](https://github.com/celebi-pkg/flight-analysis) by Kaya Celebi.

